---
title: "Comparing_datasets_by_BP_MF_CC"
author: "Vitalii Kleshchevnikov"
date: "10 May 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = FALSE, message = FALSE, fig.keep = "all") 
suppressPackageStartupMessages({
  library(data.table)
  library(dplyr)
  library(downloader)
  library(UniProt.ws)
  library(PSICQUIC)
  library(ggplot2)
  library(clusterProfiler)
})
  source("multiplot.R")
  source("GO_analyse_simplify_plot.R")
  source("isoform_id_all_remover.R")
```


```{r swissprot}
swissprot_proteins = fread("http://www.uniprot.org/uniprot/?query=reviewed:yes+AND+organism:9606&columns=id&format=tab", stringsAsFactors = F, colClasses = "character")$Entry
```

```{r IntAct}
# IntAct
all.IntAct = fread("./Data/IntAct_MITAB27_release_05-2017.txt", stringsAsFactors = F, colClasses = "character")
# changing column names to data.table-compatible format
{
(colnames_IntAct = colnames(all.IntAct))
colnames(all.IntAct) = gsub(" ","_",colnames(all.IntAct))
colnames(all.IntAct) = gsub("\\(|\\)","",colnames(all.IntAct))
colnames(all.IntAct) = gsub("#","",colnames(all.IntAct))
}
# cleaning Taxid "taxid:9606(human)|taxid:9606(Homo sapiens)" to 9606
all.IntAct = unique(all.IntAct[,.(IDs_interactor_A, IDs_interactor_B,  Taxid_interactor_A, Taxid_interactor_B, Host_organisms, Publication_Identifiers)])
{
all.IntAct[, Taxid_interactor_A := gsub("taxid:|\\(.*$","",Taxid_interactor_A)]
all.IntAct[, Taxid_interactor_B := gsub("taxid:|\\(.*$","",Taxid_interactor_B)]
all.IntAct[, Host_organisms := gsub("taxid:|\\(.*$","",Host_organisms)]
# saving identifier types and cleaning interactor ids
all.IntAct[, interactor_IDs_databases_A := gsub(":.*$","",IDs_interactor_A)]
all.IntAct[, interactor_IDs_databases_B := gsub(":.*$","",IDs_interactor_B)]
all.IntAct[, IDs_interactor_A := gsub("^.*:","",IDs_interactor_A)]
all.IntAct[, IDs_interactor_B := gsub("^.*:","",IDs_interactor_B)]
# cleaning other information
all.IntAct[, Publication_Identifiers := gsub("^.*pubmed:|\\|.*$","",Publication_Identifiers)]
all.IntAct[interactor_IDs_databases_A == "uniprotkb", IDs_interactor_A := isoform_id_all_remover(IDs_interactor_A)]
all.IntAct[interactor_IDs_databases_B == "uniprotkb", IDs_interactor_B := isoform_id_all_remover(IDs_interactor_B)]
}

all.IntAct_proteins = all.IntAct[Taxid_interactor_A == "9606" & Taxid_interactor_B == "9606" &
                                   interactor_IDs_databases_A == "uniprotkb" & interactor_IDs_databases_B == "uniprotkb" &
                                   IDs_interactor_A %in% swissprot_proteins & IDs_interactor_B %in% swissprot_proteins,
                                 unique(c(IDs_interactor_A, IDs_interactor_B))]
not_all.IntAct_proteins = swissprot_proteins[!swissprot_proteins %in% all.IntAct_proteins]
```

```{r BioGrid}
BioGrid = fread("https://raw.githubusercontent.com/vitkl/darkspaceproject/master/BioGRID/results/pairs_pmids_biogrid.txt", stringsAsFactors = F, colClasses = "character")
BioGrid[, c("ida", "idb") := tstrsplit(pair_id_clean, "_")]
BioGrid_proteins = BioGrid[ida %in% swissprot_proteins & idb %in% swissprot_proteins,
                                 unique(c(ida, idb))]
not_BioGrid_proteins = swissprot_proteins[!swissprot_proteins %in% BioGrid_proteins]
```

```{r String}
String = fread("https://raw.githubusercontent.com/vitkl/darkspaceproject/master/STRING/results/pairs_STRING_all.txt", stringsAsFactors = F, colClasses = "character")
String_proteins = String[ida_clean %in% swissprot_proteins & idb_clean %in% swissprot_proteins,
                                 unique(c(ida_clean, idb_clean))]
not_String_proteins = swissprot_proteins[!swissprot_proteins %in% String_proteins]
```

```{r Mentha}
if(!file.exists("./Data/2017-05-08_MITAB-2.5.zip")){
        mentha_url = "http://mentha.uniroma2.it/dumps/2017-05-08_MITAB-2.5.zip"
        # httr::GET(http://mentha.uniroma2.it/doDownload.php?file=2017-05-08_MITAB-2.5.zip) - this is how I got the direct link
        downloader::download(mentha_url, destfile = "./Data/2017-05-08_MITAB-2.5.zip")
}
unzip("./Data/2017-05-08_MITAB-2.5.zip", exdir = "./Data/")
Mentha = fread("./Data/2017-05-08_MITAB-2.5", stringsAsFactors = F, colClasses = "character", header = F)
{
Mentha[, Taxid_interactor_A := gsub("taxid:|\\(.*$","",V10)]
Mentha[, Taxid_interactor_B := gsub("taxid:|\\(.*$","",V11)]
# saving identifier types and cleaning interactor ids
Mentha[, interactor_IDs_databases_A := gsub(":.*$","",V1)]
Mentha[, interactor_IDs_databases_B := gsub(":.*$","",V2)]
Mentha[, IDs_interactor_A := gsub("^.*:","",V1)]
Mentha[, IDs_interactor_B := gsub("^.*:","",V2)]
# cleaning other information
Mentha[, Publication_Identifiers := gsub("^.*pubmed:|\\|.*$","",V9)]
Mentha[interactor_IDs_databases_A == "uniprotkb", IDs_interactor_A := isoform_id_all_remover(IDs_interactor_A)]
Mentha[interactor_IDs_databases_B == "uniprotkb", IDs_interactor_B := isoform_id_all_remover(IDs_interactor_B)]
}

Mentha_proteins = Mentha[IDs_interactor_A %in% swissprot_proteins & IDs_interactor_B %in% swissprot_proteins,
                                 unique(c(IDs_interactor_A, IDs_interactor_B))]
not_Mentha_proteins = swissprot_proteins[!swissprot_proteins %in% Mentha_proteins]
```

```{r Vidal}
Vidal_proteins = all.IntAct[Taxid_interactor_A == "9606" & Taxid_interactor_B == "9606" &
                             interactor_IDs_databases_A == "uniprotkb" & interactor_IDs_databases_B == "uniprotkb" &
                             Publication_Identifiers == "25416956" &
                             IDs_interactor_A %in% swissprot_proteins & IDs_interactor_B %in% swissprot_proteins,
                           unique(c(IDs_interactor_A, IDs_interactor_B))]
not_Vidal_proteins = swissprot_proteins[!swissprot_proteins %in% Vidal_proteins]
```

```{r Mann}
Mann_proteins = all.IntAct[Taxid_interactor_A == "9606" & Taxid_interactor_B == "9606" &
                             interactor_IDs_databases_A == "uniprotkb" & interactor_IDs_databases_B == "uniprotkb" &
                             Publication_Identifiers == "26496610" &
                             IDs_interactor_A %in% swissprot_proteins & IDs_interactor_B %in% swissprot_proteins,
                           unique(c(IDs_interactor_A, IDs_interactor_B))]
not_Mann_proteins = swissprot_proteins[!swissprot_proteins %in% Mann_proteins]
```

```{r Omnipath}
interactions_url = "http://omnipathdb.org/interactions/?fields=sources&fields=references"
interactions_file = "./Data/OmniPath_interactions.txt"
if(!file.exists(interactions_file)) download(interactions_url,interactions_file)
Omnipath = fread(interactions_file, stringsAsFactors = F, colClasses = "character", header = T)
Omnipath_proteins = Omnipath[source %in% swissprot_proteins & target %in% swissprot_proteins, unique(c(source, target))]
not_Omnipath_proteins = swissprot_proteins[!swissprot_proteins %in% Omnipath_proteins]
```

## Dataset size

### proteins missing from the dataset  
`r length(not_all.IntAct_proteins)` length(not_all.IntAct_proteins)  
`r length(not_BioGrid_proteins)` length(not_BioGrid_proteins)  
`r length(not_String_proteins)` length(not_String_proteins)  
`r length(not_Mentha_proteins)` length(not_Mentha_proteins)  
`r length(not_Mann_proteins)` length(not_Mann_proteins)  
`r length(not_Vidal_proteins)` length(not_Vidal_proteins)  
`r length(not_Omnipath_proteins)` length(not_Omnipath_proteins) 

### proteins missing from the dataset  
`r length(all.IntAct_proteins)` length(all.IntAct_proteins)  
`r length(BioGrid_proteins)` length(BioGrid_proteins)  
`r length(String_proteins)` length(String_proteins)  
`r length(Mentha_proteins)` length(Mentha_proteins)  
`r length(Mann_proteins)` length(Mann_proteins)  
`r length(Vidal_proteins)` length(Vidal_proteins)  
`r length(Omnipath_proteins)` length(Omnipath_proteins)  

## Enrichment in biological process

```{r BPdropGOlevels}
drop_GO_levels_ = c(1:3,13:20)
maxSetSize_ = Inf
```

```{r BPcluster_enrichment_negative, fig.height=5, fig.width=12}
not_combined = data.table(proteins = c(not_all.IntAct_proteins, not_BioGrid_proteins, not_String_proteins, not_Mentha_proteins, not_Mann_proteins, not_Vidal_proteins, not_Omnipath_proteins),
                      resources = c(rep("IntAct", length(not_all.IntAct_proteins)),
                        rep("BioGRID", length(not_BioGrid_proteins)),
                        rep("STRING", length(not_String_proteins)),
                        rep("Mentha", length(not_Mentha_proteins)),
                        rep("Mann_dataset", length(not_Mann_proteins)),
                        rep("Vidal_dataset", length(not_Vidal_proteins)),
                        rep("Omnipath", length(not_Omnipath_proteins))))

if(!file.exists("BPnot_result")){
not_result = cluster_GO_enrich_simplify_plot_bioc(formula = proteins ~ resources,
                                              protein_groups.dt = not_combined,
                                              reference_protein_set = swissprot_proteins,
                                              identifier_type = "UNIPROT",
                                              ontology = "BP",
                                              pAdjustMethod_ = "BH",
                                              minSetSize = 40, maxSetSize = maxSetSize_,
                                              simplify_by = "p.adjust",
                                              simplify_fun = "min",
                                              similarity_calc_method = "none",
                                              similarity_cutoff = 0.7,
                                              visualize_result = "dotplot",
                                              above_corrected_pval = 0.05,
                                              plot_title = "",
                                              drop_GO_levels = drop_GO_levels_)
save(not_result, file = "BPnot_result")
}
#zz = list()
#for(i in 1:length(drop_GO_levels_)){
#      zz = c(clusterProfiler:::getGOLevel(ont = "BP", level = drop_GO_levels_[i]), zz)
#      print(drop_GO_levels_[i])
#}
```

```{r BPcluster_enrichment_positive, fig.height=5, fig.width=14}
combined = data.table(proteins = c(all.IntAct_proteins, BioGrid_proteins, String_proteins, Mentha_proteins, Mann_proteins, Vidal_proteins, Omnipath_proteins),
                      resources = c(rep("IntAct", length(all.IntAct_proteins)),
                        rep("BioGRID", length(BioGrid_proteins)),
                        rep("STRING", length(String_proteins)),
                        rep("Mentha", length(Mentha_proteins)),
                        rep("Mann_dataset", length(Mann_proteins)),
                        rep("Vidal_dataset", length(Vidal_proteins)),
                        rep("Omnipath", length(Omnipath_proteins))))

if(!file.exists("BPresult")){
result = cluster_GO_enrich_simplify_plot_bioc(formula = proteins ~ resources,
                                              protein_groups.dt = combined,
                                              reference_protein_set = swissprot_proteins,
                                              identifier_type = "UNIPROT",
                                              ontology = "BP",
                                              pAdjustMethod_ = "BH",
                                              minSetSize = 40, maxSetSize = maxSetSize_,
                                              simplify_by = "p.adjust",
                                              simplify_fun = "min",
                                              similarity_calc_method = "none",
                                              similarity_cutoff = 0.7,
                                              visualize_result = "dotplot",
                                              above_corrected_pval = 0.05,
                                              plot_title = "",
                                              drop_GO_levels = drop_GO_levels_)
save(result, file = "BPresult")
}
```

## Enrichment in molecular function

```{r MFdropGOlevels}
drop_GO_levels_ = c(1:5,13:20)
maxSetSize_ = Inf
```

```{r MFcluster_enrichment_negative, fig.height=5, fig.width=12}
if(!file.exists("MFnot_result")){
not_result = cluster_GO_enrich_simplify_plot_bioc(formula = proteins ~ resources,
                                              protein_groups.dt = not_combined,
                                              reference_protein_set = swissprot_proteins,
                                              identifier_type = "UNIPROT",
                                              ontology = "MF",
                                              pAdjustMethod_ = "BH",
                                              minSetSize = 40, maxSetSize = maxSetSize_,
                                              simplify_by = "p.adjust",
                                              simplify_fun = "min",
                                              similarity_calc_method = "none",
                                              similarity_cutoff = 0.7,
                                              visualize_result = "dotplot",
                                              above_corrected_pval = 0.05,
                                              plot_title = "",
                                              drop_GO_levels = drop_GO_levels_)
save(not_result, file = "MFnot_result")
}
```

```{r MFcluster_enrichment_positive, fig.height=5, fig.width=14}
if(!file.exists("MFresult")){
result = cluster_GO_enrich_simplify_plot_bioc(formula = proteins ~ resources,
                                              protein_groups.dt = combined,
                                              reference_protein_set = swissprot_proteins,
                                              identifier_type = "UNIPROT",
                                              ontology = "MF",
                                              pAdjustMethod_ = "BH",
                                              minSetSize = 40, maxSetSize = maxSetSize_,
                                              simplify_by = "p.adjust",
                                              simplify_fun = "min",
                                              similarity_calc_method = "none",
                                              similarity_cutoff = 0.7,
                                              visualize_result = "dotplot",
                                              above_corrected_pval = 0.05,
                                              plot_title = "",
                                              drop_GO_levels = drop_GO_levels_)
save(result, file = "MFresult")
}
```

## Enrichment in cellular component

```{r CCdropGOlevels}
drop_GO_levels_ = c(1:2,7:20)
maxSetSize_ = Inf
```

```{r CCcluster_enrichment_negative, fig.height=5, fig.width=12}
if(!file.exists("CCnot_result")){
not_result = cluster_GO_enrich_simplify_plot_bioc(formula = proteins ~ resources,
                                              protein_groups.dt = not_combined,
                                              reference_protein_set = swissprot_proteins,
                                              identifier_type = "UNIPROT",
                                              ontology = "CC",
                                              pAdjustMethod_ = "BH",
                                              minSetSize = 40, maxSetSize = maxSetSize_,
                                              simplify_by = "p.adjust",
                                              simplify_fun = "min",
                                              similarity_calc_method = "none",
                                              similarity_cutoff = 0.7,
                                              visualize_result = "dotplot",
                                              above_corrected_pval = 0.05,
                                              plot_title = "",
                                              drop_GO_levels = drop_GO_levels_)
save(not_result, file = "CCnot_result")
}
```

```{r CCcluster_enrichment_positive, fig.height=5, fig.width=14}
if(!file.exists("CCresult")){
result = cluster_GO_enrich_simplify_plot_bioc(formula = proteins ~ resources,
                                              protein_groups.dt = combined,
                                              reference_protein_set = swissprot_proteins,
                                              identifier_type = "UNIPROT",
                                              ontology = "CC",
                                              pAdjustMethod_ = "BH",
                                              minSetSize = 40, maxSetSize = maxSetSize_,
                                              simplify_by = "p.adjust",
                                              simplify_fun = "min",
                                              similarity_calc_method = "none",
                                              similarity_cutoff = 0.7,
                                              visualize_result = "dotplot",
                                              above_corrected_pval = 0.05,
                                              plot_title = "",
                                              drop_GO_levels = drop_GO_levels_)
save(result, file = "CCresult")
}
```

```{r multiplot, fig.height=30, fig.width=14}
load("MFnot_result")
load("MFresult")
load("BPnot_result")
load("BPresult")
load("CCnot_result")
load("CCresult")
multiplot(plotlist = list(MFnot_result + ggtitle("under-represented molecular functions"),
                          MFresult + ggtitle("over-represented molecular functions"),
                          BPnot_result + ggtitle("under-represented biological processes"),
                          BPresult + ggtitle("over-represented biological processes"),
                          CCnot_result + ggtitle("under-represented cellular components"),
                          CCresult + ggtitle("over-represented cellular components")), cols=1)

```