---
title: "How does the use of protein as a bait depend on the protein length?"
author: "Vitalii Kleshchevnikov"
date: "13 December 2016"
output: html_document
---

### setting up parameters of the analysis
##=========================================================##
Species to select from:
"Homo sapiens", "Mus musculus", "strain ATCC 204508", "strain K12"
Paste value in the code chunk below
``` {r SPECIES_NAME} 
SPECIES_NAME = "Homo sapiens"
```
all Uniprot if reviewed == 1, only Swissprot data if reviewed == 2, 
TrEMBL data if reviewed == 3
only reviewed = 2 is relevant for this analysis 
Paste value in the code chunk below
``` {r reviewed}
reviewed = 2
```
To distinguish between isoforms or not (to use only generic UniprotACs): TRUE / FALSE?
only isoforms = FALSE is relevant for this analysis 
Paste value in the code chunk below
``` {r isoforms}
isoforms = FALSE
```
Please specify the date for which you want to perform the analysis (if not today)
"Logic table" from protein_properties script is needed for this analysis. Filename example:  
"proteome_vs_interactome_protein_properties_f_Homo sapiens_reviewed_2_isoforms_FALSE_2016-12-01.txt"
``` {r date}
date = Sys.Date()
date = as.Date("2016-12-01")
```
the date of analysis, how old is Uniprot protein list: `r date` 

### Species ID lookup
##=========================================================##
looking up species ID in the Uniprot readme file (always reads from url)
``` {r species_id_lookup}
source("SPECIES_NAME_TO_ID.R")
r = reviewed
i = isoforms
n = SPECIES_NAME
SPECIES_IDs = SPECIES_NAME_TO_ID(n)
SPECIES_ID = SPECIES_IDs$SPECIES_ID
```

### querying for interations detected by AP-MS or two hybrid using PSICQUIC 
##=========================================================##
Detection method explanation:

- detmethod:MI:0232 - transcriptional complementation assay, includes different two-hybrids
http://www.ebi.ac.uk/ols/ontologies/mi/terms?iri=http%3A%2F%2Fpurl.obolibrary.org%2Fobo%2FMI_0232
- detmethod:MI:0090 - protein complementation assay, includes transcriptional 
complementation assay and bimolecular fluorescence complementation, enzyme complementations
http://www.ebi.ac.uk/ols/ontologies/mi/terms?iri=http%3A%2F%2Fpurl.obolibrary.org%2Fobo%2FMI_0090

- both detmethod and pmethod are required for specifying AP-MS
- detmethod:MI:0004 - affinity chromatography technology, co-IP and pulldowns
http://www.ebi.ac.uk/ols/ontologies/mi/terms?iri=http%3A%2F%2Fpurl.obolibrary.org%2Fobo%2FMI_0004
- pmethod:MI:0433 - partial identification of protein sequence
http://www.ebi.ac.uk/ols/ontologies/mi/terms?iri=http%3A%2F%2Fpurl.obolibrary.org%2Fobo%2FMI_0433

to change: type in exact database names (the list below - all IMEx -is default for the function)
``` {r querying_PSICQUIC}
databases <- c("IntAct", "MINT", "bhf-ucl", "MPIDB", "MatrixDB", 
               "HPIDb","I2D-IMEx","InnateDB-IMEx", "MolCon", "UniProt", "MBInfo")

source("query_PSICQUIC_for_interactions.R")
all_interactions = query_PSICQUIC_for_interactions(SPECIES_ID = SPECIES_ID, 
                                                   SPECIES_NAME = SPECIES_NAME, 
                                                   databases = databases, date)
twohybrids_all_interactions = query_PSICQUIC_for_interactions(SPECIES_ID = SPECIES_ID, 
                                                   SPECIES_NAME = SPECIES_NAME, 
                                                   databases = databases, date,
                                                   detmethod = "transcriptional complementation assay")
complementation_all_interactions = query_PSICQUIC_for_interactions(SPECIES_ID = SPECIES_ID, 
                                                              SPECIES_NAME = SPECIES_NAME, 
                                                              databases = databases, date,
                                                              detmethod = "protein complementation assay")
ap_ms_all_interactions = query_PSICQUIC_for_interactions(SPECIES_ID = SPECIES_ID, 
                                                         SPECIES_NAME = SPECIES_NAME, 
                                                         databases = databases, date,
                                                         detmethod = "affinity chromatography technology",
                                                         pmethod = "partial identification of protein sequence")
```
### Cleaning and transforming data
##=========================================================##
spliting all interactions into baits and preys
```{r baits_and_preys}
all_interactions.bait
twohybrids_all_interactions.bait
complementation_all_interactions.bait
ap_ms_all_interactions.bait
all_interactions.prey
twohybrids_all_interactions.prey
complementation_all_interactions.prey
ap_ms_all_interactions.prey
```
extracting interactor ID-s from interactions (MI-TAB 2.5)
```{r interactions_to_interactors}
source("interactions_to_interactors.R")
twohybrids_all_interactors = interactions_to_interactors(twohybrids_all_interactions)
complementation_all_interactors = interactions_to_interactors(complementation_all_interactions)
ap_ms_all_interactors = interactions_to_interactors(ap_ms_all_interactions)
```
filtering interactors for uniprotkb only indentifiers 
filtering for SPECIES_ID only proteins
```{r uniprotkb_and_SPECIES_ID_interactor_selector}
source("uniprotkb_and_SPECIES_ID_interactor_selector.R")
twohybrids_all_interactors_SPECIES_ID_only = uniprotkb_and_SPECIES_ID_interactor_selector(twohybrids_all_interactors, SPECIES_ID)
complementation_all_interactors_SPECIES_ID_only = uniprotkb_and_SPECIES_ID_interactor_selector(complementation_all_interactors, SPECIES_ID)
ap_ms_all_interactors_SPECIES_ID_only = uniprotkb_and_SPECIES_ID_interactor_selector(ap_ms_all_interactors, SPECIES_ID)
```
Removing all isoform IDs (XXXXXX-X+ => XXXXXX) from IDs
``` {r isoform_id_all_remover}
  source("isoform_id_all_remover.R")
  twohybrids_all_interactors_SPECIES_ID_only$interactor_IDs = isoform_id_all_remover(twohybrids_all_interactors_SPECIES_ID_only$interactor_IDs)
  complementation_all_interactors_SPECIES_ID_only$interactor_IDs = isoform_id_all_remover(complementation_all_interactors_SPECIES_ID_only$interactor_IDs)
  ap_ms_all_interactors_SPECIES_ID_only$interactor_IDs = isoform_id_all_remover(ap_ms_all_interactors_SPECIES_ID_only$interactor_IDs)
```
preparing interaction detection method data for logic table: selecting unique proteins and adding the column of ones
``` {r pre_logic_table}
unique_twohybrids_interactors_SPECIES_ID_only = unique(cbind(twohybrids_all_interactors_SPECIES_ID_only[c("interactor_IDs")], 1))
colnames(unique_twohybrids_interactors_SPECIES_ID_only)[2] = "two_hybrid"
unique_complementation_interactors_SPECIES_ID_only = unique(cbind(complementation_all_interactors_SPECIES_ID_only[c("interactor_IDs")], 1))
colnames(unique_complementation_interactors_SPECIES_ID_only)[2] = "all_protein_complementation"
unique_ap_ms_interactors_SPECIES_ID_only = unique(cbind(ap_ms_all_interactors_SPECIES_ID_only[c("interactor_IDs")], 1))
colnames(unique_ap_ms_interactors_SPECIES_ID_only)[2] = "AP_MS"
```
loading logic table made by "swissprot_vs_imex_protein_properties" script
```{r loading_logic_table}
filename_vs_2 = paste("./analysis/","proteome_vs_interactome_protein_properties_f_", n,"_reviewed_",r,"_isoforms_",i,"_", date,".txt", sep = "")
proteome_vs_imex_details_f = as.data.frame(read.delim(filename_vs_2, header = T, stringsAsFactors = F,quote=""))
proteome_vs_imex_details_f$whole_proteome_Uniprot_IMEx = factor(proteome_vs_imex_details_f$whole_proteome_Uniprot_IMEx, ordered =F)
```
merging new results with the logic table
```{r merging_logic_table}
proteome_vs_imex_interaction_details_t1 = merge(proteome_vs_imex_details_f, 
                                   unique_twohybrids_interactors_SPECIES_ID_only, 
                                   by.x = "whole_proteome_IDs",
                                   by.y = "interactor_IDs",
                                   all.x = T, all.y = F)
proteome_vs_imex_interaction_details_t2 = merge(proteome_vs_imex_interaction_details_t1, 
                                               unique_complementation_interactors_SPECIES_ID_only, 
                                               by.x = "whole_proteome_IDs",
                                               by.y = "interactor_IDs",
                                               all.x = T, all.y = F)
proteome_vs_imex_interaction_details_f = merge(proteome_vs_imex_interaction_details_t2, 
                                               unique_ap_ms_interactors_SPECIES_ID_only, 
                                                by.x = "whole_proteome_IDs",
                                                by.y = "interactor_IDs",
                                                all.x = T, all.y = F)
proteome_vs_imex_interaction_details_f[is.na(proteome_vs_imex_interaction_details_f)] = 0
```
adding factor combination variable - two_hybrid and AP_MS
``` {r two_hybrid_vs_AP_MS}
proteome_vs_imex_interaction_details_f[,length(proteome_vs_imex_interaction_details_f)+1] = interaction(proteome_vs_imex_interaction_details_f$two_hybrid, proteome_vs_imex_interaction_details_f$AP_MS, sep = "_")
colnames(proteome_vs_imex_interaction_details_f)[length(proteome_vs_imex_interaction_details_f)] = paste0("two_hybrid", "_vs_","AP_MS")
levels(proteome_vs_imex_interaction_details_f$two_hybrid_vs_AP_MS) = c("not_two_hybrid_and_not_AP_MS", "two_hybrid_not_AP_MS","not_two_hybrid_but_AP_MS", "two_hybrid_and_AP_MS")
proteome_vs_imex_interaction_details_f$two_hybrid_vs_AP_MS = as.character(proteome_vs_imex_interaction_details_f$two_hybrid_vs_AP_MS)
proteome_vs_imex_interaction_details_f$two_hybrid_vs_AP_MS[proteome_vs_imex_interaction_details_f$IMEx!=1] = "not_in_IMEx"
```
saving combined logic table with protein properties from Uniprot and interaction properties combined
``` {r saving_logic_table}
filename = paste("./analysis/","proteome_vs_interactome_interaction_properties_f_", n,"_reviewed_",r,"_isoforms_",i,"_", date,".txt", sep = "")
write.table(proteome_vs_imex_interaction_details_f,filename,col.names=T,row.names=F,sep="\t",quote=F)
```

# Results
##=========================================================##
read the table saved before 
- script can be started from here
``` {r reading_table}
filename = paste("./analysis/","proteome_vs_interactome_interaction_properties_f_", n,"_reviewed_",r,"_isoforms_",i,"_", date,".txt", sep = "")
proteome_vs_imex_interaction_details_f = as.data.frame(read.delim(filename, header = T, stringsAsFactors = F,quote=""))
proteome_vs_imex_interaction_details_f$whole_proteome_Uniprot_IMEx = factor(proteome_vs_imex_interaction_details_f$whole_proteome_Uniprot_IMEx, ordered =F)
proteome_vs_imex_interaction_details_f$two_hybrid_vs_AP_MS = factor(proteome_vs_imex_interaction_details_f$two_hybrid_vs_AP_MS, ordered =F)
```

### The density and the histogram of the protein mass (overlay for different detection methods)
##=========================================================##
``` {r histogram_and_density, fig.width = 10, fig.height = 7}
library(ggplot2)
ggplot(proteome_vs_imex_interaction_details_f, aes(x = Mass, color =two_hybrid_vs_AP_MS, 
                                                   fill = two_hybrid_vs_AP_MS)) +
  geom_density(alpha =0.2) + scale_x_log10() +
  xlab("protein mass, Da, log10 scale")+
  ggtitle("density plot of the protein mass (overlay for different detection methods)") #+facet_grid(two_hybrid_vs_AP_MS~Organism)
ggplot(proteome_vs_imex_interaction_details_f, aes(x = Mass,color = two_hybrid_vs_AP_MS)) +
  scale_x_log10() + geom_histogram(position = "identity", bins = 50,alpha =0.1) +
  xlab("protein mass, Da, log10 scale") + 
  ggtitle("histogram of the protein mass (overlay for different detection methods)")
#+facet_grid(two_hybrid_vs_AP_MS~Organism)
```

### How does protein interaction detection method depend on the protein length? - Violin plot
``` {r violin_plot, fig.width = 12, fig.height = 12}
library(dplyr)
library(scales) 
## adding two_hybrid and AP_MS levels to show non-combinatory relationship (not 
## excluding proteins present in both)
two_hybrid = filter(proteome_vs_imex_interaction_details_f, two_hybrid == 1)
two_hybrid$two_hybrid_vs_AP_MS = "two_hybrid"
AP_MS = filter(proteome_vs_imex_interaction_details_f, AP_MS == 1)
AP_MS$two_hybrid_vs_AP_MS = "AP_MS"
for_plot = rbind(proteome_vs_imex_interaction_details_f, two_hybrid, AP_MS)
for_plot$two_hybrid_vs_AP_MS = factor(for_plot$two_hybrid_vs_AP_MS, ordered = F)
## Calculating protein median protein mass for each group
yy = split(for_plot, for_plot$two_hybrid_vs_AP_MS)
Mass_median = sapply(yy, function(x){median(log10(x$Mass))})
kDa_median_mass = (10^Mass_median)/1000
plot_labels = paste0(gsub("_"," ",levels(for_plot$two_hybrid_vs_AP_MS)),", \n (median mass: ", signif(kDa_median_mass,3)," kDa)")
## Violin plot - proportional
ggplot(for_plot, aes(y = Mass, x =two_hybrid_vs_AP_MS, fill = two_hybrid_vs_AP_MS)) +
  geom_violin(draw_quantiles = c(0.05,0.25,0.495,0.5,0.505,0.75,0.95),scale = "count", alpha =0.7) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x))) +
  scale_x_discrete(labels = plot_labels)+
  geom_abline(slope = 0, intercept = Mass_median, alpha =0.1) +
  ylab("Mass, Da, log10 scale")+
  xlab("presence in IMEx, detection method")+
  ggtitle("How does protein interaction detection method depend on the protein length?",
          subtitle = "violin areas are scaled proportionally to the number of observations")+
  coord_flip()#+facet_grid(two_hybrid_vs_AP_MS~Organism)

## Violin plot - not proportional
ggplot(for_plot, aes(y = Mass, x =two_hybrid_vs_AP_MS, fill = two_hybrid_vs_AP_MS)) +
  geom_violin(draw_quantiles = c(0.05,0.25,0.495,0.5,0.505,0.75,0.95),scale = "area", alpha =0.7) +
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x))) +
  scale_x_discrete(labels = plot_labels)+
  geom_abline(slope = 0, intercept = Mass_median, alpha =0.1) +
  ylab("Mass, Da, log10 scale")+
  xlab("presence in IMEx, detection method")+
  ggtitle("How does protein interaction detection method depend on the protein length?",
          subtitle = "all violins have the same area")+
  coord_flip()#+facet_grid(two_hybrid_vs_AP_MS~Organism)
print(signif(data.frame("mass_median(log10.Da)" = Mass_median, "mass_median(kDa)" = (10^Mass_median)/1000),3))
```
